---
title: "Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(FitLM)
library(car)
library(bench)
```

## Background

### Linear regression

In statistics, [linear regression](https://en.wikipedia.org/wiki/Linear_regression) is a linear approach for modelling the relationship between a scalar response and one or more explanatory variables (also known as dependent and independent variables). In linear regression, the relationships are modeled using linear predictor functions whose unknown model parameters are estimated from the data.

### "lm" function in R

The existing `stats::lm` in `R` is a commonly used function that can help fit linear regression model. It can provide us with estimation, related statistics, and hypothesis test results. 

However, the `stats::lm` function can not directly output some important results,including the confidence interval(CI) of $\hat{\beta}$, general linear hypothesis(GLH) test results, partial F test results and hat matrix. 

## Basic usage

### installation

* `FitLM` can be downloaded from github:

```
devtools::install_github("fxgstuart/BIOSTAT625_HW3_temp", build_vignettes = T)
```
* You might also need to install the `devtools` first. You can configure the environment by following these steps:
 
  (1) Install the release version of devtools from CRAN with `install.packages("devtools")`.
  
  (2) Setting up working development environment:
      
      *- Windows*: Install [Rtools](https://cran.r-project.org/bin/windows/Rtools/).
      
      *- Mac*: Install [Xcode](https://developer.apple.com/xcode/) from the Mac App Store.
      
      *- Linux*: Install a compiler and various development libraries (details vary across differnet flavors of Linux).
  
  (3) Follow the instructions below depending on platform:
      
      *- Mac & Linux:* 
      ```
      devtools::install_github("hadley/devtools")
      ```
      
      *- Windows:* 
      
      ```
      library(devtools)
      build_github_devtools()
      #### Restart R before continuing ####
      install.packages("devtools.zip", repos = NULL)
      # Remove the package after installation
      unlink("devtools.zip")
      ```
      
  Find more tutorials [here](https://www.r-project.org/nosvn/pandoc/devtools.html).

### Loading packages

```
library(FitLM)
```

### Basic command

To fit the linear regression model using 'FitLM' package, generally two input needs to be prepared 

* **y**: A column vector, treated as outcome.

* **X**: A numerical matrix, treated as predictor.

The package also offers 'intercept' as a choice

* **intercept**:  Logical values, TRUE (by default) or FALSE. If "TRUE", the model will include the intercept.

Some other parameters are needed for certain functions:

For ‘fit.confint’, parameter 'level' is needed:

* **level**: numerical value, stands for the significance of the confidence interval

For 'fit.partial.test', 'idx' is required:

* **idx**: a vector of indices of tested parameters

Last but not least, for 'fit.GLH.test', 'T.matrix' and 'c' are required:

* **T.matrix**: hypothesis matrix (or vector) giving linear combinations of coefficients by rows
* **c**: right-hand-side vector for hypothesis, with as many entries as rows in the hypothesis matrix; can be omitted, in which case it defaults to a vector of zeroes

*see the help page:* (take fit.GLH.test as an example)

```
help("fit.GLH.test")
```

**Now, let's take a look at some examples of how to use the this package:**

We use the `"mtcars"` dataset as an example. `"mtcars"` is a dataframe with 32 rows and 11 numerical columns, which was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

The following statistics is provided in the respective columns:

* mpg: Miles/(US) gallon

* cyl: Number of cylinders

* disp: Displacement (cu.in.)

* hp: Gross horsepower

* drat: Rear axle ratio

* wt: Weight (1000 lbs)

* qsec: 1/4 mile time

* vs: Engine (0 = V-shaped, 1 = straight)

* am:	Transmission (0 = automatic, 1 = manual)

* gear:	Number of forward gears

* carb:	Number of carburetors

The `"mtcars"` dataset has been included in `graphics` package, which is included in `R` by default.

*(1) Fit a simple linear regression model and calculate the coefficients:*

```{r, collapse=TRUE}
y <- as.matrix(mtcars$hp)
colnames(y) <- "hp"
X <- as.matrix(mtcars$mpg)
colnames(X) <- "mpg"
SLR.model.coefs <- fit.coef(X,y)
print(SLR.model.coefs)
```

*(2) Fit a multiple linear regression model with or without an intercept:*

Regression model with the intercept:

```{r,collapse=TRUE}
y <- as.matrix(mtcars$hp)
colnames(y) <- "hp"
X <- as.matrix(mtcars[,c(1,3,5,6,7)])
MLR.model.with.intercept.coefs <- fit.coef(X, y, intercept = T)
print(MLR.model.with.intercept.coefs)
```

Regression model without the intercept:

```{r}
MLR.model.no.intercept.coefs <- fit.coef(X, y, intercept = F)
print(MLR.model.no.intercept.coefs)
```

*(4) Calculate $R^2$ and adjusted $R^2$ of linear model:*

```{r}
res <- fit.R.squared(X,y)
R.squared <- res$R.squared
adj.R.squared <- res$adj.R.squared
```

*(5) Obtain $hat$-matrix:*

```{r}
hat.matrix <- fit.hat.matrix(X)
```

*(6) Conduct General Linear Hypothesis testing:*

```{r}
contrast.matrix <- matrix(c(0, 1,-1, 0, 0, 0), nrow = 1)  # define the T matrix
c.matrix <- c(0)  # define the constant matrix
res <- fit.GLH.test(X,y,T.matrix = contrast.matrix, c = c.matrix)
print(res)
``` 

## Correctness

**(1) Compare the point estimate and standard error of $\beta$'s, and results from the partial $T$-test**
```{r,collapse=TRUE}
y <- as.matrix(mtcars$hp)
X <- as.matrix(mtcars[,c(1,3,5,6,7)])
res.fit.coef <- fit.coef(X,y)
res.fit.sd <- fit.sd(X,y)
res.fit.t.test <- fit.t.test(X,y)
res.tb <- cbind(res.fit.coef,res.fit.sd, res.fit.t.test$t, res.fit.t.test$p.value)
colnames(res.tb) <- c("Estimation","Standard error","t","p-value")
print(res.tb)
```

For comparison, we use `stats::lm()` function to fit this model to prove the correctness of our package:

```{r,collapse=TRUE}
summary(lm(y ~ X))$coefficients
```

```{r,collapse=TRUE}
all.equal(unname(res.tb), 
          unname(summary(lm(y ~ X))$coefficients))
```

**(2) Calculate the 95% confidence interval of $\beta$'s:**

```{r, collapse=TRUE}
CI.95.actual <- fit.confint(X,y,level = 0.95)
print(CI.95.actual)
```

The above results are are consistent with the results from `stats::confint`: 

```{r}
CI.95.expected <- confint(lm(y ~ X))
print(CI.95.expected)
```

```{r}
all.equal(unname(CI.95.actual), unname(CI.95.expected))
```

**(3) Results of overall F-test:**

```{r, collapse=TRUE}
actual.F.test <- fit.overall.test(X,y)
expected.F.test <- as.numeric(summary(lm(y ~ X))$fstatistic)
all.equal(unname(unlist(actual.F.test[c(1,3,4)])), expected.F.test)  #compare overall f-statistics
```

**(4) Results of $R^2$ and adjusted $R^2$:**

```{r, collapse=TRUE}
res.R.squared <- fit.R.squared(X,y)
actual.R.Squared <- as.numeric(res.R.squared$R.squared) 
expected.R.Squared <- as.numeric(summary(lm(y ~ X))$r.squared)
all.equal(actual.R.Squared, expected.R.Squared)
actual.adjusted.R.Squared <- as.numeric(res.R.squared$adj.R.squared)
expected.adjusted.R.Squared <- as.numeric(summary(lm(y ~ X))$adj.r.squared)
all.equal(actual.adjusted.R.Squared, expected.adjusted.R.Squared)
```

**(5) Conduct General Linear Hypothesis testing:**

```{r}
contrast.matrix <- matrix(c(0, 1,-1, 0, 0, 0), nrow = 1)  # define the contrast matrix
c.matrix <- c(0)  # define the constant matrix
res.GLH <- fit.GLH.test(X,y,T.matrix = contrast.matrix, c = c.matrix)
``` 

We use `car::linearHypothesis()` function to conduct the GLH test, and compare the results with those from `linear_Regression()`:

```{r}
GLH.expected <- car::linearHypothesis(model = lm(y ~ X), 
                                      hypothesis.matrix = contrast.matrix, rhs = c.matrix)
all.equal(unname(unlist(res.GLH)), as.vector(c(GLH.expected$F[2], GLH.expected$`Pr(>F)`[2])))
```

## Efficiency

Let's compare the speed between `FitLM` and `stats::lm` by using `bench:mark()` function:

*(Note: package `bench` should be installed)*

```{r,collapse=TRUE}
y <- as.matrix(mtcars$hp)
X <- as.matrix(mtcars[,c(1,3,5,6,7)])
code_efficiency_comparison <-  bench::mark("FitLM" = {
  beta.actual <- fit.coef(X,y)
  as.vector(beta.actual)
}, "stat::lm" = {
  beta.expected <- lm(y ~ X)$coefficients
  as.vector(beta.expected)
})  
summary(code_efficiency_comparison)
itr.per.sec <- as.numeric(code_efficiency_comparison[[4]])
itr.per.sec.ratio <- itr.per.sec[1] / itr.per.sec[2]
```

From the results of `bench::mark`, it can be seen that `FitLM` is about `r round(itr.per.sec.ratio, digits = 2)` times as faster as `stats:lm`.

Furthermore, we can compare the speed of constructing confidence interval:

```{r}
code_efficiency_comparison_CI <-  bench::mark("FitLM" = {
  CI.actual <- fit.confint(X,y)
  unname(as.matrix(CI.actual))
}, "stat::lm" = {
  CI.expected <- confint(lm(y~X))
  unname(as.matrix(CI.expected))
})  
summary(code_efficiency_comparison_CI)
itr.per.sec <- as.numeric(code_efficiency_comparison_CI[[4]])
itr.per.sec.ratio <- itr.per.sec[1] / itr.per.sec[2]
```

From the above results, it can be seen that `FitLM::fit.confint` is about `r round(itr.per.sec.ratio, digits = 2)` times as faster as `stats:lm`.

Also, we compare the speed of conducting GLH test:

```{r}
contrast.matrix <- matrix(c(0, 1,-1, 0, 0, 0), nrow = 1)  
c.matrix <- c(0)  
code_efficiency_comparison.GLH <- bench::mark("FitLM" = {
  GLH.F.statistics.actual <- fit.GLH.test(X,y,T.matrix = contrast.matrix, c = c.matrix)
  as.vector(GLH.F.statistics.actual$F.stat)
}, "car::linearHypothesis" = {
  GLH.F.statistics.expected <- 
    car::linearHypothesis(model = lm(y ~ X), 
                          hypothesis.matrix = contrast.matrix, rhs = c.matrix)$F[2]
  as.vector(GLH.F.statistics.expected)
})
summary(code_efficiency_comparison.GLH)
GLH.itr.per.sec <- as.numeric(code_efficiency_comparison.GLH[[4]])
GLH.itr.per.sec.ratio <- GLH.itr.per.sec[1] / GLH.itr.per.sec[2]
```

It can be seen that `FitLM::fit.GLH.test` is about `r round(GLH.itr.per.sec.ratio, digits = 2)` times as faster as `car::linearHypothesis` when conducting the GLH test.
